---
title: 'Reproducible Research: Peer Assessment 1'
author: "Zoey Le"
date: "February 4, 2018"
output: html_document
keep_md: true
---
### Setup to show every code chunks
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data  
Look at data summary and few first lines

```{r loading}
data=read.csv("activity.csv")
data$date=as.Date(as.character(data$date))
summary(data)
head(data)
```
## What is mean total number of steps taken per day?

**1. Total steps taken on each day**
```{r stepbyday}
byStep=aggregate(data$steps,by=list(data$date),sum)
names(byStep)=c("Date","TotalSteps")
head(byStep)
```

**2. Histogram of the total number of steps taken each day**
```{r histday}
hist(byStep$TotalSteps,breaks = 25,col = "steelblue",border = "white",
     main = "Total number of Steps per Day", xlab = "Total Steps",ylim = c(0,20))
```

**3. Mean and median of the total number of steps taken per day**

```{r meanmedian}
mean=round(mean(byStep$TotalSteps,na.rm=TRUE),2)
median=round(median(byStep$TotalSteps,na.rm = TRUE),2)
```

- Mean of the total number of steps taken per day is: `r toString(mean)`
- Median of the total number of steps taken per day is: `r toString(median)`

## What is the average daily activity pattern?

**1. Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**

```{r timeseries}
byInterval=aggregate(na.omit(data)$steps,by=list(na.omit(data)$interval),mean)
names(byInterval)=c("Interval","Mean.Steps")

plot(x=byInterval$Interval,y=byInterval$Mean.Steps,type = "l",col =10,lwd=2,main ="Average of Steps accross all Days",xlab="Interval",ylab = "Step")
```

**2. The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps**

```{r max}
byInterval[byInterval$Mean.Steps==max(byInterval$Mean.Steps),]
```

## Imputing missing values
**1. Total number of missing values in the dataset (i.e. the total number of rows with NAs)**

```{r NAs}
sapply(X=data,FUN = function(x) sum(is.na(x)))
```

**2. All the NAs will be replaced by the mean of steps taken on its interval**

**3. New dataset that is equal to the original dataset but with the missing data filled in.**

```{r completedata}
 complete.data=data
for (i in c(which(is.na(data$steps)))) {
  complete.data$steps[i]=byInterval[which(byInterval$Interval==data$interval[i]),2]
}
 sapply(X=complete.data,FUN = function(x) sum(is.na(x)))
```

**4. Histogram of the total number of steps taken each day**

```{r histnew}
byStep.new=aggregate(complete.data$steps,by=list(complete.data$date),sum)
names(byStep.new)=c("Date","TotalSteps")
hist(byStep.new$TotalSteps,breaks = 25,col = "steelblue",border = "white",
     main = "Total number of Steps per Day", xlab = "Total Steps",
     ylim = c(0,20))
```

**5. Any differences in mean and median between original data and data with NAs replaced?**  

Compare between histogram

```{r newmeanmedian}
newmean=round(mean(byStep.new$TotalSteps),2)
newmedian=round(median(byStep.new$TotalSteps),2)
par(mfrow=c(1,2))
hist(byStep$TotalSteps,breaks = 25,col = "steelblue",border = "white",
     main=NULL,xlab = "Total Steps per day before fill NAs",ylim = c(0,20))
hist(byStep.new$TotalSteps,breaks = 25,col = "steelblue",border = "white",
     main=NULL,xlab = "Total Steps per day after fill NAs",
     ylim = c(0,20))
```

- New Mean of the total number of steps taken per day is: `r toString(newmean)`
- New median of the total number of steps taken per day is: `r toString(newmedian)`
- The mean between those two data is the same which is `r toString(mean)`. However, the median rise from `r toString(median)` to `r toString(newmedian)`

## Are there differences in activity patterns between weekdays and weekends?

**1. New factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day**

```{r dataday}
new.data=complete.data
new.data$day.of.week=ifelse(weekdays(as.Date(new.data$date)) %in% c("Saturday","Sunday"),
                     "weekend","weekday")
new.data=aggregate(new.data$steps,by=list(new.data$interval,new.data$day.of.week),mean)
names(new.data)=c("Interval","day.of.week","Mean.Steps")
head(new.data)
```

**2. Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)**

```{r timeseriesday}
library(ggplot2)
ggplot(new.data,aes(x=Interval,y=Mean.Steps,color=day.of.week))+
  geom_line()+
  facet_grid(day.of.week~.)+
  labs(title="Means of Steps by Interval", x="Interval",y="Mean Steps")
```
